Fractional anisotropy as a tracer of cosmic voids

### 2014-09-24 cosmic_voids.tex
Deleting information of catalog of halos

At present, studying voids may be considered as a threefold enterprise
\citep{Platen07}: first, they are a key ingredient of the Cosmic Web as 
they dominate almost the entire volume distribution at large-scales and 
additionally, they compensate overdense structures in the total budget of 
matter. This implies that a full understanding of their evolution and 
properties is essential for fathoming the high complexity of the Cosmic 
Web. Second, they provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. Third, they constitute and unique and still 
largely pristine environment in which can be tested galaxy evolution.

@@

Following the recent trend of using digital image-processing techniques 
developed in other disciplines (e.g. medical sciences and computer imaging) 
for analysing the structure of the Cosmic Web, we propose here, much in 
the same way as \citet{Libeskind13}, the fractional anisotropy (FA) as a 
tracer of the internal structure and the outline of cosmic voids. The FA 
was initially introduced by \citet{Basser95} for quantifying the 
anisotropy degree of the diffusivity of water molecules through cerebral
tissue in nuclear magnetic resonance imaging. In this context we use the 
same quantity for determining the anisotropy degree of the local 
environment from either the orbital dynamics as set by the eigenvalues of
the tidal tensor (T-web), or the dynamics of the velocity field as set by 
the eigenvalues of the shear tensor (V-web). In both cases, the FA is not
determined directly from the density field, hence it is suitable for 
describing both, the structure of high non-linear regions (e.g. filaments,
knots and very dense walls) as well as for revealing the fainter 
substructure exhibited by quasi-linear regions like voids.

Once established the FA as the tracer field of cosmic voids, we proceed to
identify individual voids as basins of local minima. For this purpose we
implement a technique defined by \citet{Beucher79} and \citet{Beucher93} 
and widely used in the field of image analysis and mathematical morphology, 
i.e. the \textit{watershed transform algorithm}. This technique has been 
already implemented on void finding schemes by \citet{Platen07} and 
\citet{Neyrinck08} with very interesting results, where individual voids 
are identified as catching basins of local minima of the density field. 
The appeal of this algorithm lies in that is parameter-free and does not 
require to make any assumption on the shape and morphology of voids. 
Although we use a \textit{cloud-in-cell} (\texttt{CIC}) algorithm on a 
Cartesian mesh for estimating the density and tensor fields, instead of 
the more sophisticated \textit{Delaunay tessellation for field estimator} 
(\texttt{DTFE}) technique \citep{Schaap00}, our implementation of the 
watershed transform should not be significantly affected as we are 
interested in quasi-linear regions where the \texttt{CIC} gives similar 
estimations.

This paper is organized as follows. In section 2 we describe the used
simulation for testing our void finding algorithm, i.e. the Bolshoi
simulation. Alternatively to the web schemes T-web and V-web, we propose 
a very basic web classifier based on three threshold values defined on the
density field, this is used in our comparisons as a representative of all 
density-based schemes. A detailed description of these schemes is 
presented in section 3. In section 4 we describe how the FA can be used 
for tracing the structure of voids. Once obtained meaningful catalogues of 
voids for each scheme, we proceed to compare some typical and general 
interest properties of voids such as volume functions, distributions of 
size and density and velocity profiles. This is done in section 5. Finally
in section 6 we analyse and evaluate our findings.

@@

With the above parameters the mass of each particle in the simulation is 
$m_{\rm p}=1.4\times 10^{8}$\hMsun. Throughout this paper we shall use a
catalogue of halos identified through the Bound Density Maximum algorithm
\citep{Klypin99}. These halos have been selected to have an overdensity of
200 times the critical density in addition to be mutually excluded of the
virial radius of other halos. Mass, comoving position and peculiar 
velocity of each halo at $z=0$ were obtained through the public available 
Multidark database \footnote{\url{http://www.multidark.org/MultiDark/}} 
\citep{Riebe11}. For estimating the density and velocity fields we use a 
\textit{cloud-in-cell} (\texttt{CIC}) algorithm onto a grid of $256^3$ 
cells, corresponding to a resolution of $0.98$\hMpc per cell side. Then, 
through finite-differences and FFT methods the tidal and shear tensors 
are computed. Finally the eigenvalues and eigenvectors of each tensor are
obtained for each cell of the grid. Neglecting substructures presented 
below Megaparsec scales and taking into account our focus in voids, which
are a prominent characteristic of the Megaparsec Universe, we apply a 
Gaussian softening of one cell to all fields.

@@

This scheme was initially proposed by \citet{Hahn07} as a novel 
alternative for classifying the Cosmic Web based on the tidal tensor, 
that is somehow more fundamental than the density field as it also allows 
to quantify the orbital dynamics of the matter field. This approach is 
reached by a second-order expansion of the equations of motion around 
local minima of the gravitational potential and then extended to any 
position. The second-order term corresponds to the tidal tensor, which is
defined as the Hessian of the normalized gravitational potential

@@

In next section we shall describe an alternative approximation to select 
the threshold value $\lambda_{th}$, and as will be shown, one of its main 
advantages is the similar volume filling fraction (VFF) of each 
environment as compared between both web schemes. Using this result, we 
propose here a heuristic classification of the Cosmic Web based on 
three density thresholds assigned to each boundary (i.e. void-sheet, 
sheet-filament and filament-knot)\footnote{The progressive sequence 
void-sheet-filament-knot along any path in the Cosmic Web should be always
satisfied as each used field is smoothed.} such that the corresponding VFF 
values are similar to the values obtained for the T-web and the V-web, 
specifically they are defined to be the arithmetic mean of both schemes.

@@

According to the recent growing interest in studying galaxy formation in 
low-density regions as cosmological tests, classifying void regions is 
becoming an important task in cosmology. Most of those classification 
schemes for voids in cosmological simulations are based upon the density 
field, setting a cut off value below which some region becomes a void 
\SRKED{[references]}. Some more advanced classification schemes are based 
on Voronoi tessellations applied over the tracer particles of the 
simulation in order to compute the density field. Then, through a watershed 
transform, a hierarchy of void regions are found \SRKED{[references, 
ZOBOV algorithm]}.

As has been established \SRKED{[references]}, both web schemes presented 
in the previous section (V-web and T-web) for classifying the cosmic web 
present many advantages compared with classification schemes based
completely upon the density field, e.g. a more robust description of the
dynamic a kinematic of the cosmic web, a more reliable quantification of 
the visual impression, among others. With the aim of exploiting all of 
these advantages, we propose here a novel approach to classify voids in
cosmological simulations based entirely on the web schemes.

The original version of the T-web scheme \SRKED{[reference, Hahn]} was not
successful at reproducing the visual impression of the cosmic web, 
however, with the introduction of a threshold parameter \SRKED{[reference, 
Forero-Romero]}, this scheme, and even the V-web \SRKED{[reference, 
Hoffman]}, improved enormously. As this free parameter controls the visual 
impression provided by each scheme, phenomenons like percolation depends 
on it as well. Although percolation is one of the key features of the 
structure of void regions, indicating how voids are merged among them, and 
how they permeate all the cosmic web, our primal interest here is studying 
properties of single voids. Nevertheless, in the next section we shall 
analyse briefly the percolation phenomenon for both web schemes.

In order to deal with percolation of voids in our classification scheme,
we introduce the fractional anisotropy as defined in \SRKED{[reference, 
Libeskind]}.

%.........................................................................
%Fractional anisotropy
\eq{fractional_anisotropy}
{ FA = \frac{1}{\sqrt{3}}\sqrt{ \frac{ (\lambda_1 - \lambda_3)^2 + 
(\lambda_2 - \lambda_3)^2 + (\lambda_1 - \lambda_2)^2}{ \lambda_1^2 + 
\lambda_2^2 + \lambda_3^2} } }
%.........................................................................
where the eigenvalues are taken from any of the two web schemes. This 
index, such as it is defined, allows quantifying the local anisotropy 
degree of the cosmological environment, where $FA=0$ corresponds to highly
isotropic regions and $FA=1$ anisotropic ones.

>>>>

At present, studying voids may be considered as a threefold enterprise
\citep{Platen07}: first, they are a key ingredient of the Cosmic Web as 
they dominate almost the entire volume distribution at large-scales and 
additionally, they compensate overdense structures in the total budget of 
matter. This implies that a full understanding of their evolution and 
properties is essential for fathoming the high complexity of the Cosmic 
Web. Second, they provide a valuable resource for inferring and probing 
cosmological parameters as their structure and dynamics are highly 
determined by these values. Third, they constitute an unique and still 
largely pristine environment in which can be tested galaxy evolution.

@@

Following the recent trend of using digital image-processing techniques 
developed in other disciplines (e.g. medical sciences and computer imaging) 
for analysing the structure of the Cosmic Web, we propose here, much in 
the same way as \citet{Libeskind13}, the fractional anisotropy (FA) as a 
tracer of the internal structure and the outline of cosmic voids. The FA 
was initially introduced by \citet{Basser95} for quantifying the 
anisotropy degree of the diffusivity of water molecules through cerebral
tissue in nuclear magnetic resonance imaging. In this context we use the 
same quantity for determining the anisotropy degree of the local 
environment from either the orbital dynamics as set by the eigenvalues of
the tidal tensor (T-web), or the dynamics of the velocity field as set by 
the eigenvalues of the shear tensor (V-web). In both cases, the FA is not
determined directly from the density field, it hence is suitable for 
describing both, the structure of high non-linear regions (e.g. filaments,
knots and very dense walls) as well as the fainter substructure exhibited 
by quasi-linear regions like voids.

Once established the FA as the tracer field of cosmic voids, we proceed to
identify individual voids as basins of local minima. For this purpose we
implement a technique developed by \citet{Beucher79} and \citet{Beucher93} 
and widely used in the field of image analysis and mathematical morphology, 
i.e. the \textit{watershed transform algorithm}. This technique has been 
already implemented on void finding schemes by \citet{Platen07} and 
\citet{Neyrinck08} with very interesting results, where individual voids 
are identified as catching basins of local minima of the density field. 
The appeal of this algorithm consists in that is parameter-free and does 
not require any assumption on the shape and morphology of voids. 
Although we use a \textit{cloud-in-cell} (\texttt{CIC}) algorithm on a 
Cartesian mesh for estimating the density and tensor fields, instead of 
the more sophisticated \textit{Delaunay tessellation for field estimator} 
(\texttt{DTFE}) technique \citep{Schaap00}, our implementation of the 
watershed transform should not be significantly affected as we are 
interested in quasi-linear regions where the \texttt{CIC} gives similar 
estimations.

This paper is organized as follows. In section 2 we describe the used
simulation for testing our void finding algorithm, i.e. the Bolshoi
simulation. Alternatively to the web schemes T-web and V-web, we propose 
a heuristic web classifier based on three thresholds of the density field. 
This is used in order to illustrate the differences of our scheme with a
density-based classifier. A detailed description of these methods is 
presented in section 3. In section 4 we describe how the FA can be used 
for tracing the structure of voids. Once obtained meaningful catalogues 
of voids for each scheme, we proceed to compare some typical and general 
interest properties of voids such as volume functions, distributions of 
size and density and velocity profiles. This is done in section 5. Finally 
in section 6 we analyse and evaluate our findings.

@@

For estimating the density and velocity fields we use a 
\textit{cloud-in-cell} (\texttt{CIC}) algorithm onto a grid of $256^3$ 
cells, corresponding to a resolution of $0.98$\hMpc per cell side. Then, 
through finite-differences and FFT methods the tidal and shear tensors 
are computed. Finally, the eigenvalues and eigenvectors of the tensor are
obtained for each cell of the grid. Neglecting substructures presented 
below Megaparsec scales and taking into account our focus in voids, which
are a prominent characteristic of the Megaparsec Universe, we apply a 
Gaussian softening of one cell to all fields.

@@

This scheme was initially proposed by \citet{Hahn07} as a novel 
alternative for classifying the Cosmic Web based on the tidal tensor, 
that is somehow more fundamental than the density field as it also allows 
to quantify the orbital dynamics of the matter field. This approach 
consists of a second-order expansion of the equations of motion around 
local minima of the gravitational potential and then extended to any 
position. The second-order term corresponds to the tidal tensor, which is
defined as the Hessian matrix of the normalized gravitational potential

@@

In next section we shall describe an alternative approximation to select 
the threshold value $\lambda_{th}$. One of the main results is the similar 
volume filling fraction (VFF) for each environment in both web schemes. 
Using this result, we propose here a heuristic classification of the 
Cosmic Web based on three density thresholds assigned to each boundary 
(i.e. void-sheet, sheet-filament and filament-knot)\footnote{The 
progressive sequence void-sheet-filament-knot along any path through the 
Cosmic Web should be always satisfied as each used field is smoothed.} 
such that the corresponding VFF values are similar to the values obtained 
for the T-web and the V-web, specifically they are defined to be the 
arithmetic mean of both schemes.

@@

The fractional anisotropy (FA), as developed by \citet{Basser95}, was 
conceived to quantify the anisotropy degree of a diffusion process, e.g. 
the diffusivity of water molecules through cerebral issue in nuclear
magnetic resonance imaging. Here we propose the FA, much in the same 
way as \citet{Libeskind13}, as a tracer of cosmic voids.

%.........................................................................
%Fractional anisotropy
\eq{fractional_anisotropy}
{ FA = \frac{1}{\sqrt{3}}\sqrt{ \frac{ (\lambda_1 - \lambda_3)^2 + 
(\lambda_2 - \lambda_3)^2 + (\lambda_1 - \lambda_2)^2}{ \lambda_1^2 + 
\lambda_2^2 + \lambda_3^2} } }
%.........................................................................
where the eigenvalues are taken from either the T-web or the V-web 
(FA-Tweb and FA-Vweb respectively). Such as it is defined, $FA=0$ 
corresponds with a isotropic dynamic and $FA=1$ with a highly anisotropic 
distribution.
